---
title: "QC of samples"
author: "Anna V. Protasio"
date: "03/10/2019"
output: html_document
---

## Load samples 

Back at processing the peaks, let's load DiffBind, plus other packages that we need to execute many functions in this experiment. 

```{r, setup, warning=FALSE, message=FALSE}
require("knitr")
# specified the root directory for the project and is maintain throughout Notebook
knitr::opts_chunk$set(echo = F, warning=FALSE, message=FALSE, root.dir = "/Volumes/GoogleDrive/Shared drives/Douse_Protasio/git_repo/cut_and_run_douse")

library(DiffBind)
library(GenomicRanges)
library(DESeq2)
library(gdata)
library(rprojroot)

register(SerialParam()) 
```

Load data sheet with samples:

```{r}
samples <- read.csv(file = "data/metadata/metadata-avp25-20191003.csv",  header = T)
samples$Treatment = factor(c(rep(c(rep("ctrl",2),rep("exp",2)),3)))
metadata <- dba(sampleSheet = samples)
```

## QC samples, PCA. 

Distance matrix of samples. Notice how the samples group my replicate.

```{r}
plot(metadata)
```

PCA of samples - just another way of displaying te same results.

```{r}
dba.plotPCA(metadata, label = DBA_REPLICATE)
```

There is quite a lot of noise in the samples, likley arising from the fact that each group of replicates where sequencd using different machines (different technologies?) and these were performed at different times. 

```{r}
save.image("RData/00_QC.RData")
```

