---
title: "Finding consensus peaks"
author: "Anna V. Protasio"
date: "02/10/2019"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---


Load packages.

```{r, setup, message=}
library(DiffBind)
library(GenomicRanges)
library(DESeq2)
library(gdata)
library(GWASTools)

register(SerialParam())

# specified the root directory for the project and is maintain throughout Notebook
require(knitr)
opts_knit$set(root.dir = "/Volumes/GoogleDrive/Shared drives/Douse_Protasio/git_repo/cut_and_run_douse", echo = T, message = F, include = T)

```

Load data from QC step:

```{r}
load("RData/00_QC.RData")
```

Because we are expecting TASOR *TKO* and *mutT* samples to loose peaks rather then gain, we will derive a consensus peak set from the control samples *D3* and *WTT*. 

```{r make_consensus_peaks}
# retrieve consensus peaks from control samples
ctrl_samples = samples[which(samples$Treatment == "ctrl"),]
ctrl_mdata = dba(sampleSheet = ctrl_samples )
ctrl_cons_pks <- dba.peakset(ctrl_mdata, minOverlap=3, bRetrieve=TRUE)
```

`ctrl_cons_pks` is GRanges object with `r length(ctrl_cons_pks)` rows.

The next step is to filter out all peaks that fall in telomeres or centromeres. 

```{r remove_telo_centro}

telo = rtracklayer::import("data/annotation/telomeres_hg38.bed", format = "BED")

# load cetromeres
data(centromeres.hg38)
# change the denomination of chromosomes
centromeres.hg38$chrom = paste("chr",centromeres.hg38$chrom, sep = "")
# make it s Grange
cent = makeGRangesFromDataFrame(centromeres.hg38, ignore.strand=T, start.field="left.base",end.field="right.base")

# produce the union of both granges
excl = union(telo, cent)
saveRDS(excl, file = "RData/telomere_and_centromere.RData")

# exclude the telomeric and centromeric regions
ctrl_cons_pks_clean = setdiff(ctrl_cons_pks, excl)
rtracklayer::export(ctrl_cons_pks_clean, "output/D3_WTT_cons_peaks.bed", format = "BED")
```

This operation leave us with a reduction from `r length(ctrl_cons_pks)` to `r length(ctrl_cons_pks_clean)`.

The consensus peak set has been written to `output/D3_WTT_cons_peaks.bed`. 

```{r wrap-up}
save.image("RData/01_find_consensus_peaks.RData")
```


<!-- #### TKO vs. ctrl - experiment 1 -->

<!-- We now create a dataset that includes only the controls D3 and WTT and one experiment TKO and load the consensus peaks. The table with samples and peaks to analyse looks like: -->

<!-- ```{r tko_vs_ctrl} -->
<!-- exp1 = c("D3","TKO","WTT") -->
<!-- exp1_samples = samples[which(samples$Condition %in% exp1),] -->
<!-- exp1_mdata = dba(sampleSheet = exp1_samples) -->
<!-- # load peakset -->
<!-- exp1_mdata = dba.peakset(exp1_mdata, peaks = ctrl_cons_pks) -->
<!-- kable(dba.show(exp1_mdata)) -->
<!-- ``` -->

<!-- Where the last line represents the consensus peak set. -->

<!-- ```{r} -->
<!-- dba.overlap(exp1_mdata, exp1_mdata$masks$exp, mode=DBA_OLAP_RATE) -->
<!-- dba.plotVenn(exp1_mdata, exp1_mdata$masks$exp) -->

<!-- exp1_count = dba.count(exp1_mdata, summits = median(width(ctrl_cons_pks))) -->



<!-- ``` -->







